First, I would like to thank my advisers Dan Connors and Manish Vachharajani 
for their advice, tutoring, and incredible patience. I am grateful for the courses they 
taught which got me interested in computer engineering and lured me away
from control systems which was my initial idea of a master's degree.

Second, I'd like to thank Professor Dirk Grunwald for his motivation and 
suggestions every time I stumbled into his office without an appointment
which shows the level of his patience and the willingness to help.  

Third, I'd like to thank Tipp Moseley a fellow graduate student for helping me
out during my initial introduction into kernel hacking, and answering 
my questions whenever I felt that I was somehow within the grasp of a black hole. 
This does not exclude the very helpful and friendly Linux kernel community 
who helped a novice transition into a real kernel hacker. This would not end 
without thanking Linus Torvalds for creating the Linux kernel in the first place.

Forth, I would like to thank my fiance Sushma for listening to me whine during 
the initial part of this thesis when whatever I touched seemed to crash; For 
listening to me gloat in pride when things worked for a change, and finally
for proof reading my thesis when I realized that I have forgotten the entire
English language.

Finally, I'd like to thank my parents for motivating me all along, my friends 
for helping in proof reading my thesis document and answering any questions I had
along the way. 